---
title: "Datathon Dec2014 - Election Canada"
author: "Data4Good"
date: '2014-12-27'
output: html_document
---
<style type="text/css">
	.DPBox {
		color:#000000;
		background-color:#A0E0E0;
		border:#668B8B 1px solid;
		border-radius:6px;
		padding:10px;
	}
	
	.QBox {
		color:#000000;
		background-color:#FFB6C1;
		border:#EE6363 1px solid;
		border-radius:6px;
		padding:10px;
	}
	
	.TextWarning { color:#FF0000; font-weight:bold;}

	#kab1 { border-collapse:collapse; color:#000000; }
	#kab1 > caption { color:#0022CC; font-weight:bold; text-align:left; }
	#kab1 th { background:#444444; color:#FFFFFF; font-weight:normal; }
	#kab1 td, th { padding:3px; border:#A0A0A0 1px solid; }
	#kab1 tr:nth-child(odd)  { background: #EEEEEE; }
	#kab1 tr:nth-child(even) { background: #D8D8D8; }

	#kab2
	{ 
		border-collapse:collapse;
		border-spacing:0;
		font-family: Calibri, Arial, Helvetica, sans-serif;
		font-size:10px;
		color:#000000;
	}
	#kab2 > caption { color:#0022CC; font-size:13px; font-weight:bold; text-align:left; }
	#kab2 th { background:#666666; color:#FFFFFF; font-weight:normal; }
	#kab2 td, th { padding:1px 4px; border: 1px solid transparent; }
	#kab2 tr:nth-child(odd)  { background: #EEEEEE; }
	#kab2 tr:nth-child(even) { background: #D8D8D8; }
	/*
	#kab2 thead, th { background:#604A33; color:#FFFFFF; font-weight:normal; }
	#kab2 tr:nth-child(odd)  { background: #FFF2E2; }
	#kab2 tr:nth-child(even) { background: #E0D0C2; }
	*/
</style>
<br/>

## Synopsis

On the [Datathon](http://www.meetup.com/DataforGood/events/218482002/) organized by [Data4Good](http://www.meetup.com/DataforGood/) in December 2014, the theme was to analyze the data of Election Canada related to donations to various political parties. This is the analysis results of our team:

- [Tri Nguyen](mailto:tritanix@gmail.com?Subject=Datathon%20Election%20Canada)
- Terry Lin
- Mohamed Ismail
- Wesley Li
- Elise Nguyen


## Source Data

The Dataset contains the records of donations to 5 major political parties across all Canada covering the period 2004 - 2013:

- Original data at [Elections Canada](http://www.elections.ca/WPAPPS/WPF/EN/EDA/SelectAssociations?act=C2&part=OBS&returntype=1&option=11)

- Pre-processed data by [https://github.com/leonL/federal-contributions-munging](https://github.com/leonL/federal-contributions-munging/tree/master/munged_data)

<br/>

The pre-processed data has done a major cleanup of the orginal data. The data we use for this analysis is the `all_contributions_2004_to_2013.csv` file generated by the script [concatenate_all_csvs.R](https://github.com/leonL/federal-contributions-munging/blob/master/munged_data/concatenate_all_csvs.R) on 2014-12-17. For convenience, the CSV file is supplied in [bz2 compressed format](../all_contributions_2004_to_2013.csv.bz2) along with this document. 
The CSV file still suffers some minor data quality issues, but is largely usable to conduct the analysis:

<br/>

| Info |  Description |
|-------|----------|
| CSV Size | 283 MB (uncompressed size, All provinces, period 2004 - 2013) |
| Total records | 1,633,883 |
| Province | 4572 Unknown (0.3%). After extra cleanup, it remains 1958 (0.1%) unknown province codes, accounting for $540K donation |
| Cities | 1419 missing city names |
| . | We have made an extra cleanup and have been able to recover most of the missing province codes and to fix the names for the most important cities |
| Contributor "Home" Riding Names | 92,541 missing (5.6%) |
| "Target" Riding Names| 1,452,448 missing (88.8%) |
| . | The riding names are severely missing. The analysis results related to Ridings are mostly inaccurate. However, the analysis code is generic and can be reapplied at a later time on higher quality data.|
| contributor_id | 240,235 unique contributors, 0 missing. This is just to emphasize that the **contributor_id** is critical to the accuracy of this analysis. Because almost every result in this analysis is based on various counts of contributors. Care must be taken to determine this ID in the pre-processed data |


```{r loadData, echo=FALSE, message=FALSE, cache=TRUE, results='hide'}
library(data.table)
dtEC <- fread("~/Documents/RProjects/zMisc/ElectionCanada_LeonL/munged_data/all_contributions_2004_to_2013.csv", header=TRUE, sep = ",")

#-------------------------------------------------------------
# Extra CleanUp
#-------------------------------------------------------------

# Only keep records having contrib > 0
dtEC <- dtEC[!is.na(contribution_amount) & contribution_amount > 0]

# convert string date to Date
dtEC[, contribution_date.adjusted := as.Date(contribution_date.adjusted, "%Y-%m-%d")]

# create new column "Year"
dtEC[, Year := factor(format(contribution_date.adjusted, "%Y"))]

# convert amount from Cents to Dollars
dtEC[, contribution_amount := contribution_amount/100]


#-------------------------------------------------------------
# Province Code CleanUp
#-------------------------------------------------------------

CANprovinces = c("AB", "BC", "MB", "NB", "NL", "NS", "NT", "NU", "ON", "PE", "QC", "SK", "YT")

library(magrittr)
dtEC[, province := province %>%	
	gsub("[^a-zA-Z]", "", .) %>% #remove non-alpha
	gsub("[é]", "e", .) %>% #convert accent
	toupper %>%		
	gsub("^QU.*" , "QC", .) %>%
	gsub("^QQC.*", "QC", .) %>%
	gsub("^PQ.*" , "QC", .) %>%
	gsub("^ON.*" , "ON", .) %>%
	gsub("^OON.*", "ON", .) %>%
	gsub("^AL.*" , "AB", .) %>%
	gsub("^AB.*" , "AB", .) %>%
	gsub("^MAN.*", "MB", .) %>%
	gsub("MA|MT" , "MB", .) %>%
	gsub("^NS.*" , "NS", .) %>%
	gsub("^NOVA.*", "NS", .) %>%
	gsub("^PEI.*", "PE", .) %>%
	gsub("^SAS.*", "SK", .) %>%
	gsub("^S.*"  , "SK", .) %>%
	gsub("YK"    , "YT", .)
]

# There are probably +accurate ways to fix incorrect Province Code
# for the purpose of this analyze we assign to a generic code
dtEC[! province %in% CANprovinces, province := "zOTHER"]
dtEC[, province := factor(province)]


#-------------------------------------------------------------
# City name cleanup
#-------------------------------------------------------------

dtEC[, city := city %>%
	gsub("&apos;", "", ., ignore.case=T) %>%
	toupper %>%
	gsub("Montr[ée]al", "MONTREAL", ., ignore.case=T) %>%
	gsub("Qu[ée]bec", "QUEBEC", ., ignore.case=T) %>%
	gsub("ST\\.* JOHN'S", "ST JOHN'S", ., ignore.case=T) %>%
	gsub("IQALUIT|Iqualuit", "IQALUIT", ., ignore.case=T)
]

```

## Analysis Results

<div class="QBox">**1. Tax Credits**<br/>How much public money pumped into the system each year</div>
<br/>

[Tax Calculation Rules](http://www.elections.ca/content.aspx?section=res&dir=ces&document=part6&lang=e),
Section "Tax Credits for Political Contributions"

```{r TaxCredits, echo=FALSE, message=FALSE, cache=TRUE, results='asis'}
#-------------------------------------------------------------
# Tax Credit Calculation (by Elise Nguyen)
# TEST:
#   DonatedAmt <- c(1000, 1200, 5000, 10000, 40000, 75000, 100000, 1e6)
#   data.frame(DonatedAmt, TaxCredits = sapply(DonatedAmt, TaxCreditCalc))
#-------------------------------------------------------------
TaxCreditCalc <- function(amountDonated) # amountDonated per person per year
{
	if (amountDonated <= 40000) { 
		credits <- amountDonated * 0.75
	} else if (amountDonated <= 75000) {
		credits <- 40000 * 0.75 + (amountDonated - 40000)* 0.50    
	} else {
		credits <- 40000 * 0.75 + 35000 * 0.50 + min((amountDonated - 75000) * 1/3, 65000)
	}	
	return(credits)
} 

dtTaxCredits <- dtEC[, .(Year, contributor_id, contribution_amount)
	] [, .(ContribAmt = as.integer(sum(contribution_amount))), by=list(Year, contributor_id)
	] [, TaxCredits := sapply(ContribAmt, TaxCreditCalc)
	] [, .(YearAmt = sum(ContribAmt)/1E6, YearTaxCredits = sum(TaxCredits)/1E6), by=Year]

dtSummaryTax <- dtTaxCredits[, .(Year,
			'YearAmt (million)' = round(YearAmt,2), 
			'YearTaxCredits (million)' = round(YearTaxCredits,2),
			Ratio = paste0(round(100 * YearTaxCredits / YearAmt,0), "%")) ] %>%
	rbind(dtTaxCredits[, .(Year="TOTAL",
			'YearAmt (million)' = round(sum(YearAmt),2),
			'YearTaxCredits (million)' = round(sum(YearTaxCredits),2),
			Ratio = paste0(round(100 * sum(YearTaxCredits) / sum(YearAmt),0), "%")) ])

knitr::kable(dtSummaryTax, format="html", table.attr="id=\"kab1\"", row.names=FALSE,
				 caption = "Contribution vs Tax Credits")
```

<br/>

```{r TaxPlot, echo=FALSE, message=FALSE, fig.width=10}
library(reshape2)
library(ggplot2)

ggplot(melt(dtTaxCredits, id="Year", variable.name="AmountType"),
		 aes(x=Year, y=value, fill=AmountType)) +
	geom_bar(stat="identity", position="dodge", alpha=0.6) +
	labs(title = "Tax Credits vs Contribution", x = "", y = "Million $ CAD") +
	scale_fill_manual(values=c("dodgerblue4", "green4"), labels=c("Total Contribution", "Tax Credits")) +
	guides(fill=guide_legend(title=NULL)) + # Legends: remove title
	theme(legend.position = c(0,1), legend.justification = c(0,1),
			legend.text = element_text(face="bold", size=10),
			legend.background = element_rect(fill=FALSE, color="black", linetype=0)) +	
	theme(axis.text.x = element_text(color="black", face="bold"))
```

<div class="QBox">**2A. Regional Variation**<br/>Highest Contributing Cities</div>
<br/>

```{r Top50Canada, echo=FALSE, message=FALSE, cache=TRUE, results='asis'}
dtCityRank <- dtEC[, .(ContribCount=.N, AmountKCAD=round(sum(contribution_amount)/1000,0)), by=list(city, province)
	] [ order(-ContribCount)] [, RankByCount := seq_len(.N)
	] [ order(-AmountKCAD)] [, RankByAmount := seq_len(.N)
	] [, .(RankByAmount, city, province, AmountKCAD, ContribCount, RankByCount,
			 AvgAmtByContrib = paste0("$", round(1000 * AmountKCAD/ContribCount, 0))) ]

knitr::kable(dtCityRank[1:50,], format="html", table.attr="id=\"kab2\"", row.names=FALSE,
				 caption="Top50 Contributing Cities in Canada (2004-2013)")
```

<br/><br/>

```{r TopNProvince, echo=FALSE, message=FALSE, cache=TRUE, results='asis'}
dtTopNCities <- dtEC[, .(province, city, contribution_amount)
	] [, .(ContribCount=.N, AmountKCAD = round(sum(contribution_amount)/1000, 0)), by=list(province, city)
	] [ order(province, -AmountKCAD)
	] [, .SD[c(1:5)], by=province # select Top N rows, groupby province
	] [ order(province, -AmountKCAD)
	] [, RankByAmount := seq_len(.N), by=province
	] [, .(province, city, ContribCount, AmountKCAD,
			 AvgAmtByContrib = paste0("$", round(1000 * AmountKCAD/ContribCount, 0)))]

knitr::kable(dtTopNCities, format="html", table.attr="id=\"kab2\"", row.names=TRUE,
				 caption="Top 5 Cities per Province (2004-2013)")
```

<br/><br/>

```{r TopCityPlot, echo=FALSE, message=FALSE, fig.width=10, fig.height=8}
library(data.table)
library(reshape2)

dtTopCityProvAmount <- dtEC[, .(province, city, contribution_amount)
	] [, .(ContribCount=.N,
			 AmountKCAD = round(sum(contribution_amount)/1000, 0),
			 AvgAmtByContrib = round(sum(contribution_amount)/.N, 0)), by=list(province, city)
	] [ order(province, -AmountKCAD)
	] [, .SD[1], by=province # select Top1 
	] [, .(province, city, ProvCity = paste0("(", province, ") ", city), AmountKCAD, ContribCount, AvgAmtByContrib)
	] [ order(-AmountKCAD)]

topCityProvPlot <- melt(dtTopCityProvAmount[,
	.(ProvCity, 'Total Amount ($ x1000)'=AmountKCAD, 'Average $ per Donor'=AvgAmtByContrib)],
	id="ProvCity", variable.name="AmountType")

ggplot(topCityProvPlot, aes(x=reorder(ProvCity, -value), y=value, fill=AmountType, ymax=value*1.05)) +
	geom_bar(stat="identity", alpha=0.7) +
	labs(title = "Top City Donation by Amount per Province (2004-2013)", x = "", y = "") +
	theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1, color="black", size=rel(1.2), face="bold")) +
	scale_fill_manual(values=c("#553322", "#AA8877"), guide=FALSE) +
	geom_text(aes(label=ifelse(AmountType=="Total Amount ($ x1000)",
										format(value, big.mark=",", scientific=FALSE), paste0("$", value)), vjust=-0.5),
				 position = position_dodge(width=0.9), size=4, color="#553322") +
	facet_grid(AmountType ~ ., scales="free") +
	theme(strip.text = element_text(face="bold", color="white", size=rel(1.2)), strip.background=element_rect(fill="grey30", size=1))
```

<br/>
<div class="QBox">**2B. Regional Variation**<br/>Highest Contributing Province / Per Capita</div>
<br/>

Additional Data: [Population per Province](http://www12.statcan.gc.ca/census-recensement/2011/dp-pd/hlt-fst/pd-pl/Table-Tableau.cfm?LANG=Eng&T=101&S=50&O=A)

```{r ProvCapita, echo=FALSE, message=FALSE, cache=TRUE, fig.width=10}
dtPopulationProv <- data.table(
	province = CANprovinces,
	Population = c(3645257, 4400057, 1208268, 751171, 514536, 921727, 41462, 31906, 12851821, 140204, 7903001, 1033381, 33897),
	key = "province"
)

setkey(dtEC, province) # same join KEY on 2 DTs

dtContribProv <- dtEC[ , .(province, contribution_amount)
	] [dtPopulationProv, list(
		TotalProv_Dollar_x1000 = round(sum(contribution_amount)/1000, 0),
		PerCapita_Dollar = round(sum(contribution_amount)/Population, 2)),
		by=.EACHI, nomatch = 0L
	] [, .(province, PerCapita_Dollar, TotalProv_Dollar_x1000)]


ggplot(dtContribProv, aes(x=province, y=PerCapita_Dollar, ymax=max(PerCapita_Dollar)*1.05 )) +
	geom_bar(stat="identity", position="dodge", fill="deeppink4", alpha=0.6) +
	labs(y="Amount $CAD") +
	theme(plot.title=element_blank(), axis.title.x=element_blank()) +
	theme(axis.text.x = element_text(color="black", size=rel(1.2), face="bold")) +
	geom_text(aes(label=paste0("$", PerCapita_Dollar), vjust=-0.5), position = position_dodge(width=0.9), size=4) +
	scale_y_continuous(breaks=seq(0,25,5), labels=c("0", "$5", "$10", "$15", "$20", "$25")) +
	annotate("text", x=0 , y=Inf, size=6, hjust=0, vjust=2, fontface="bold", label=" Contribution per Capita (2004-2013)")

ggplot(dtContribProv, aes(x=province, y=TotalProv_Dollar_x1000, ymax=max(TotalProv_Dollar_x1000)*1.10)) +
	geom_bar(stat="identity", position="dodge", fill="dodgerblue4", alpha=0.6) +
	labs(y="Amount $CAD x1000") +
	theme(plot.title=element_blank(), axis.title.x=element_blank()) +
	theme(axis.text.x = element_text(color="black", size=rel(1.2), face="bold")) +
	geom_text(aes(label=format(TotalProv_Dollar_x1000, big.mark=",", scientific=FALSE), vjust=-0.5), position = position_dodge(width=0.9), size=4) +
	scale_y_continuous(breaks=c(1E3, 1E4, 2E4, 3E4, 4E4, 5E4, 75E3, 1E5, 130E3),
							 labels=c("1K", "10K", "20K", "30K", "40K", "50K", "75K", "100K", "130K")) +
	annotate("text", x=0 , y=Inf, size=6, hjust=0, vjust=1.5, fontface="bold", label=" Total Province Contribution (2004-2013)")
```

<br/>
<div class="QBox">**3. Trends of Donation to Political Parties per Year**</div>
<br/>

```{r TrendsParty, echo=FALSE, message=FALSE, cache=TRUE, fig.width=10}
dtPartyYear <- dtEC[ , .(Year, party_name, contribution_amount)
	] [, .(AmountKCAD = round(sum(contribution_amount)/1000, 0)),
		by=list(Year, party_name)
	] [ order(Year, party_name)]

ggplot(dtPartyYear, aes(x = Year, y=AmountKCAD, color=party_name, group=party_name)) +
	geom_point(size=3, shape=19) +
	geom_line(size=1) +
	labs(title = "Contribution per Party", x = "", y = "Million $CAD") +
	theme(axis.text.x = element_text(color="black", size=rel(1.2), face="bold")) +
	scale_color_manual(values = c("dodgerblue2", "navy", "forestgreen", "red3", "#FF7F00"), name="Party Name") +
	scale_y_continuous(breaks=c(5e3, 10e3, 15e3, 20e3), labels=c("$5M", "$10M", "$15M", "$20M"))	
```

<br/>
<div class="QBox">**4. Single or Multiple Parties?**<br/>How many people donate to 1, 2, 3, 4 Parties?</div>
<br/>

Several analysis have been done on this question. The number of persons contributing to one single party represents the overwhelming majority. The simplest results is the summary table below. The counts are made on **unique** contributor_id for the entire dataset.

Please note that it is also possible to break down the count per Year. The code is supplied as comments in the R Markdown document. But the results are not displayed here.


```{r HowManyParty, echo=FALSE, message=FALSE, cache=TRUE, results='asis'}
dtPartyFreqAll <- dtEC[, .(party_name, contributor_id),
	] [, .(DonationCount = .N), by = list(party_name, contributor_id)
	] [, .(PartyFreq = .N), by = list (contributor_id)
	] [, .(DonorCount = .N), by = list (PartyFreq)
	] [, TotalDonor := sum(DonorCount)
	] [, PartyFreq := factor(PartyFreq)
	] [ order(PartyFreq), .(PartyFreq, DonorCount, TotalDonor,
			 PercentTotal = paste0(round(100 * DonorCount/TotalDonor, 1), "%"))]

# # Counts Break down By Year
# dtPartyFreqByYear <- dtEC[, .(Year, party_name, contributor_id),
# 	] [, .(DonationCount = .N), by = list(Year, party_name, contributor_id)
# 	] [, .(PartyFreq = .N), by = list (Year, contributor_id)
# 	] [, .(DonorCount = .N), by = list (Year, PartyFreq)
# 	] [, TotalDonorYear := sum(DonorCount), by = Year
# 	] [, PartyFreq := factor(PartyFreq)
# 	] [ order(Year, PartyFreq), .(Year, PartyFreq, DonorCount, TotalDonorYear,
# 					 PercentTotal = paste0(round(100 * DonorCount/TotalDonorYear, 1), "%"))]

knitr::kable(dtPartyFreqAll, format="html", table.attr="id=\"kab1\"", row.names=FALSE,
				 caption="Number of Unique Contributors per Frequency of Parties Donated (2004 - 2013)")
```

<br/>

```{r HowManyPartyPlot, echo=FALSE, message=FALSE}
backupBaseMargins <- par("mar") # default: par(mar = c(5.1,4.1,4.1,2.1)) : c(bottom, left, top, right)
par(mar = c(1, 1, 1, 1))

pie(dtPartyFreqAll$DonorCount,
	 main = "Number of Unique Contributors Donating to N Parties (2004 - 2013)",
	 # manual labels suppress labels for PartyFreq=3, PartyFreq=4 (to avoid overlap)
	 labels = c(sprintf("PartyFreq = 1\n%s contributors\n(%s)",
			format(dtPartyFreqAll[PartyFreq==1, .(DonorCount)], big.mark=",", scientific=FALSE),
			dtPartyFreqAll[PartyFreq==1, .(PercentTotal)]),
		sprintf("PartyFreq = 2\n%s contributors\n(%s)",
			format(dtPartyFreqAll[PartyFreq==2, .(DonorCount)], big.mark=",", scientific=FALSE),
			dtPartyFreqAll[PartyFreq==2, .(PercentTotal)])),
	 col=c("paleturquoise3", "#F4A460", "purple3", "black"))

par(mar = backupBaseMargins)
```

<div class="QBox">**5A. Donation Schedule**<br/>How many people donate one lump sum every year?</div>
<br/>
A lump sum Donation is when the contributor makes ONE single donation a year.

```{r 5ALumpSum, echo=FALSE, message=FALSE, cache=TRUE, fig.width=10}
dtOnceAYear <- dtEC[ , .(Year, contributor_id)
	] [, .(DonationCount = .N), by=list(Year, contributor_id)
	] [, TotalDonor := .N, by=Year
	] [ DonationCount==1
	] [, .(LumpsumDonor=.N, Ratio = paste0(round(100 * .N / TotalDonor,0), "%")),
			by=list(Year, TotalDonor)]

ggplot(dtOnceAYear) +
	labs(title="Contributors making a Lump sum Donation per Year", x="", y="Number of Contributors") +
	theme(axis.text.x = element_text(color="black", size=rel(1.2), face="bold")) +
	geom_bar(aes(x=Year, y=TotalDonor), stat="identity", fill="grey40", alpha=0.5) +
	geom_text(aes(x=Year, y=TotalDonor, label=TotalDonor, vjust=-0.5), position = position_dodge(width=0.9), size=4) +
	geom_bar(aes(x=Year, y=LumpsumDonor), stat="identity", fill="midnightblue", alpha=0.6) +
	geom_text(aes(x=Year, y=LumpsumDonor, label=LumpsumDonor, vjust=-0.5), position = position_dodge(width=0.9), size=4, color="midnightblue") +
	geom_text(aes(x=Year, y=LumpsumDonor, label=Ratio, vjust=-2), position = position_dodge(width=0.9), size=4) +
	annotate("text", x=0.5 , y=Inf, hjust=0, vjust=1.5, fontface="bold", col="grey40", size=4, label="xxx: Total Contributors") +
	annotate("text", x=0.5 , y=Inf, hjust=0, vjust=3.5, fontface="bold", col="midnightblue", size=4, label="xxx: \"Lump sum\" Contributors")	
```

<br/>
<div class="QBox">**5B. Donation Schedule**<br/>How many donate more than once per year?</div>
<br/>
In the majority of case, the Frequency of donation varies from 1 to 12 times per year.
There are some outliers above 12 up to 155 times per year. For these outliers we group them together by range.

```{r 5BMultiDonation, echo=FALSE, message=FALSE, cache=TRUE, results='asis'}
dtMultiDonation <- dtEC[ , .(Year, contributor_id)
	] [, .(DonationCount = .N), by=list(Year, contributor_id)
	] [, Frequency := sapply(DonationCount, cut, 
				breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 15, 20, 30, Inf),
				labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13-15", "16-20", "21-30", "31+"))
	] [, .(DonorCount = .N), by = list(Year, Frequency)
	] [, TotalDonor := sum(DonorCount), by = Year
	] [, PercentTotal := round(100 * DonorCount/TotalDonor, 1)
	] [ order(Year, Frequency) ]

# xTab Showing Year & TotalDonor in rows, Donation Frequency in columns
xtabYearFreq <- reshape2::dcast(dtMultiDonation, Year + TotalDonor ~ Frequency, value.var = "DonorCount", fill=0)

knitr::kable(xtabYearFreq, format="html", table.attr="id=\"kab1\"", row.names=FALSE,
				 caption="Frequency of Donation per Year")
```

<br/><br/>
To summarize the table above visually, we average all the values for the entire period in a single plot.
Although the averaged values over 10 years are not 100% representative of a frequency within a year. In practice,
the values are quite consistent year over year. The "average" plot below is reasonably correct to show the relative proportions between various frequencies.

```{r 5BMultiDonationPlot, echo=FALSE, message=FALSE, cache=TRUE, fig.width=10}
dtPlotRecurDonor <- dtMultiDonation[, .(DonorCount = sum(DonorCount)), by=Frequency
	] [, TotalDonor := sum(DonorCount) ]

ggplot(dtPlotRecurDonor, aes(x=Frequency, y=DonorCount, fill=Frequency, ymax=max(DonorCount)*1.10)) +
	geom_bar(stat="identity", alpha=0.7) +
	labs(title="Frequency of Donation per Year\n(Average on the entire period 2004-2013)", x = "", y = "Number of Contributors") +
	theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1, color="black", size=rel(1.2), face="bold")) +
	guides(fill=FALSE) +
	geom_text(aes(label=format(DonorCount, big.mark=",", scientific=FALSE), vjust=-0.5), position = position_dodge(width=0.9), size=4) +
	geom_text(aes(label=paste0(round(100*DonorCount/TotalDonor,1), "%"), vjust=-2.2),
				 position = position_dodge(width=0.9), size=4, color="blue")
```

<br/>
<div class="QBox">**6. Riding vs Federal**<br/>How many people donating at both levels (to National Party and to the local Riding)</div>
<br/>

|-----|-----|
| **Single Level** | means the Contributor makes a donation to either at the Riding level OR at the Federal level |
| **Multi Level**  | means the Contributor makes donations to both the Riding AND Federal levels |

<br/>

```{r 6RidingAndFederal, echo=FALSE, message=FALSE, cache=TRUE, results='asis'}
# ContribCount = the count of donation per Contributor
#                At max, there are two records per Contributor,
#                1 for federal_contribution=FALSE and 1 for federal_contribution=TRUE
# LevelCount: 1 record per Contributor
#   value = 1 : means Contributor donated to 1 level (either federal only or local riding only)
#   value = 2 : means Contributor donated to both Federal and local riding

dtDonationByLevel <- dtEC[, .(Year, contributor_id, federal_contribution)
	] [, .(ContribCount = .N), by = list(Year, contributor_id, federal_contribution)
	] [, .(LevelCount = .N), by = list(Year, contributor_id)
	] [, .(SingleLevel = sum(ifelse(LevelCount == 1, 1, 0)),
			 MultiLevel  = sum(ifelse(LevelCount >  1, 1, 0)),
			 TotalDonor  = .N), by=Year
	] [, .(Year, 
		SingleLevel, Ratio1 = paste0(round(100* SingleLevel/TotalDonor, 0), "%"),
		MultiLevel , Ratio2 = paste0(round(100* MultiLevel/TotalDonor, 0), "%"), TotalDonor) ]

knitr::kable(dtDonationByLevel, format="html", table.attr="id=\"kab1\"", row.names=FALSE,
				 caption="Contribution by Level of Donation")
```

<br/><br/>

```{r 6RidingAndFederalPlot, echo=FALSE, message=FALSE, cache=TRUE, fig.width=10}
dtPlotLevelOfContrib <- melt(dtDonationByLevel[, .(Year, SingleLevel, MultiLevel)],
									  id.vars="Year", variable.name="LevelOfContrib")

ggplot(dtPlotLevelOfContrib,
		 aes(x=Year, y=value, ymax=max(value)*1.10, fill=LevelOfContrib)) +
	geom_bar(stat="identity", alpha=0.6) +
	labs(title="Levels of Contribution", x = "", y = "Number of Contributors") +
	guides(fill=guide_legend(title=NULL, reverse=TRUE)) + # Legends: remove title, reverse item order
	scale_fill_manual(values = c("#006670", "maroon4"), #name="Level of Contribution",
							labels=c("1Lev: Federal OR Riding", "2Lev: Both Federal & Riding")) +
	theme(legend.position = c(0,1), legend.justification = c(0,1),
			legend.text = element_text(face="bold", size=10),
			legend.background = element_rect(fill=FALSE, color="black", linetype=0)) +
	theme(axis.text.x = element_text(color="black", size=rel(1.2), face="bold"))

```

<br/>
<div class="QBox">**7. Home Riding vs Away Riding**<br/>Number of Donors inside and outside of the "home" Riding</div>
<br/>

Please note that the data suffers a severe quality issue on the Riding information:

- **92,541** records (5.6%) missing "Home" Riding Names
- **1,452,448** records (88.8%) missing "Target" Riding Names
- **NOTE:** Consistency naming between Home and Target ridings is critical for this question. The analysis uses the ridings names *as-is*. The results presented below is just informative and should be considered as <span class="TextWarning">**inaccurate**</span>. However, the analysis code is not data dependent and should be rerun on cleaner data (filling missing and normalized riding names).

|-----|-----|
| **SingleRiding** | number of Contributors to one single riding |
| **MultiRiding**  | number of Contributors to more than one riding |
| **MissingRiding**| number of Contributors when either home/target riding name is missing |

<br/>

```{r 7HomeAwayRiding, echo=FALSE, message=FALSE, cache=TRUE, results='asis'}
# Total records where donation was made to an Unknown riding (missing riding name)
# nrow(dtEC[is.na(contributors_riding_name)]) # [1] 92541
# nrow(dtEC[is.na(target_riding)]) # [1] 1452448

dtDonorCount <- dtEC[, .(ContribCount=.N), by=.(Year, contributor_id)] [, .(TotalUniqueDonor = .N), by=Year]

dtContribRiding <- 
dtEC[!is.na(contributors_riding_name) & !is.na(target_riding),
	] [, .(Year, contributor_id, ContribInHomeRiding = (contributors_riding_name == target_riding))
	] [, .(ContribCount = .N), by = list(Year, contributor_id, ContribInHomeRiding)
	] [, .(RidingSideCount = .N), by = list(Year, contributor_id)
	] [, .(SingleRiding = sum(ifelse(RidingSideCount == 1, 1, 0)),
			 MultiRiding  = sum(ifelse(RidingSideCount >  1, 1, 0))), by=Year]

# Count of Contributors
# ATTENTION:
# SingleRiding: nb of Donors to 1 single riding
# MultiRiding : nb of Donors to more than 1 Riding
# MissingRiding: nb of Donors when either home/target riding name is missing
setkey(dtDonorCount, Year)
setkey(dtContribRiding, Year)

dtDonationRiding <- dtContribRiding[dtDonorCount, .(Year,
	SingleRiding, PercentSingle = paste0(round(100 * SingleRiding/TotalUniqueDonor,1), "%"),
	MultiRiding , PercentMulti  = paste0(round(100 * MultiRiding/TotalUniqueDonor, 1), "%"),
	MissingRiding = TotalUniqueDonor - SingleRiding - MultiRiding,
	MissingPercent= paste0(round(100 * (1- (SingleRiding+MultiRiding)/TotalUniqueDonor), 1), "%"),
	TotalUniqueDonor), nomatch=0L]

setkey(dtDonationRiding, Year)

knitr::kable(dtDonationRiding, format="html", table.attr="id=\"kab1\"", row.names=FALSE,
				 caption="Count of Contributors")
```

<br/><br/>

```{r 7HomeAwayRidingPlot, echo=FALSE, message=FALSE, cache=TRUE, fig.width=10}
dtPlotRiding <- melt(dtDonationRiding[, .(Year, SingleRiding, MultiRiding, MissingRiding)],
									  id.vars="Year", variable.name="RidingVariety")

ggplot() +
	geom_bar(data = dtPlotRiding,
		 aes(x=Year, y=value, ymax=max(value)*1.10, fill=RidingVariety),
		 stat="identity", alpha=0.6) +
	labs(title="Contribution by Number of Ridings", x = "", y = "Number of Contributors") +
	guides(fill=guide_legend(title=NULL, reverse=TRUE)) + # Legends: remove title, reverse item order
	scale_fill_manual(values = c("seagreen", "red4", "grey40"),
							labels=c("Single Riding", "More than 1 Riding", "Riding Name Unknown")) +
	theme(legend.position = c(0,1), legend.justification = c(0,1),
			legend.text = element_text(face="bold", size=10),
			legend.background = element_rect(fill=FALSE, color="black", linetype=0)) +
	theme(axis.text.x = element_text(color="black", size=rel(1.2), face="bold")) +
	geom_text(data = dtDonationRiding[, .(Year, TotalUniqueDonor)],
				 aes(x=Year, y=TotalUniqueDonor, label=TotalUniqueDonor, vjust=-0.5), size=4)
```

## Environment

- R version 3.1.2 (2014-10-31)
- RStudio Version 0.98.1091
- Platform: x86_64-pc-linux-gnu (64-bit)

